2.3 Prob Model for PCR Positivity Estimation

### 2024-04-22 2_scripts/2.3 Prob Model for PCR pos.qmd
update figures for AABB abstract (unfinished)

## add a data source indicator
table(tb_alberta$assay, tb_alberta$raw)
tb_alberta$data[tb_alberta$assay=="prop.abbott"
                & tb_alberta$raw=="raw"]<-"APL"
tb_alberta$data[tb_alberta$assay=="prop.roche"
                & tb_alberta$raw=="transformed"]<-"APL"
tb_alberta$data[tb_alberta$assay=="prop.abbott"
                & tb_alberta$raw=="transformed"]<-"CBS"
tb_alberta$data[tb_alberta$assay=="prop.roche"
                & tb_alberta$raw=="raw"]<-"CBS"
table(tb_alberta$data, tb_alberta$assay)
table(tb_alberta$data, tb_alberta$raw)
```

Alton's comments:
- solid/dash to indicate raw/transformed
- red/blue to indicate Abbott/Roche

```{r}
# making the plot, 
p03.3<-tb_alberta %>%
  ggplot(aes(x=samplemonth, y=prop, colour=data)) +
  geom_point(aes(colour=data, shape=raw)) +
  geom_line(aes(colour=data, linetype=raw))  +
  theme_bw() + 
  xlab("Time of Sample Collection") + 
  ylab("Sero-prevalence") 
p03.3
```

@@

p05
```
#### Alton suggests using marginal dist with {ggExtra}
```{r}
ggMarginal(p04, groupColour = TRUE, groupFill = TRUE) # log-scale
ggMarginal(p05, groupColour = TRUE, groupFill = TRUE) # with CBS matched data
```

>>>>

## add a data source indicator
table(tb_alberta$assay, tb_alberta$raw)
tb_alberta$data[tb_alberta$assay=="prop.abbott"
                & tb_alberta$raw=="raw"]<-"APL"
tb_alberta$data[tb_alberta$assay=="prop.roche"
                & tb_alberta$raw=="transformed"]<-"APL"
tb_alberta$data[tb_alberta$assay=="prop.abbott"
                & tb_alberta$raw=="transformed"]<-"CBS"
tb_alberta$data[tb_alberta$assay=="prop.roche"
                & tb_alberta$raw=="raw"]<-"CBS"
table(tb_alberta$data, tb_alberta$assay)
table(tb_alberta$data, tb_alberta$raw)

## create a merged indicator for study + transformed
## we could use interaction() in ggplot instead
tb_alberta$study<-paste(tb_alberta$data, 
                        rep(" (", length(tb_alberta$data)),
                        tb_alberta$raw, 
                        rep(")", length(tb_alberta$data) ))
```

Alton's comments:
- solid/dash to indicate raw/transformed
- red/blue to indicate Abbott/Roche
- Merge the color & linetype indicator in the legend

```{r}
# making the plot, either those codes will work
p03.3<-tb_alberta %>%
  ggplot(aes(x=samplemonth, y=prop, colour=study, linetype=study)) +
  geom_point() +
  geom_line()  +
  theme_bw() + 
  xlab("Time of Sample Collection") + 
  ylab("Sero-prevalence")+ 
  scale_colour_manual("",   values = c("red", "red", "blue", "blue"))+
  scale_linetype_manual("", values=c(1,3,1,3))

# use this chunk to skip the paste()
p03.lab<-c("APL (raw)", "CBS (raw)", 
           "APL (transformed)", "CBS (transformed)")
p03.3<-tb_alberta %>%
  ggplot(aes(x=samplemonth, y=prop, colour=interaction(data, raw),
                                    linetype=interaction(data, raw))) +
  geom_point() +
  geom_line()  +
  theme_bw() +
  xlab("Time of Sample Collection") + 
  ylab("Sero-prevalence")+ 
  scale_colour_manual("", values = c("red", "blue", "red", "blue"),
                      labels = p03.lab)+
  scale_linetype_manual("", values=c(1,1,3,3), labels = p03.lab) +
  theme(legend.position="top")+
  guides(color = guide_legend(nrow = 2))

p03.3
```

@@

p05
```
#### Alton suggests using marginal dist with {ggExtra}
```{r}
ggMarginal(p04, groupColour = TRUE, groupFill = TRUE) # log-scale
ggMarginal(p05, groupColour = TRUE, groupFill = TRUE) # with CBS matched data
```
#### 3.5 Create figure panel for AABB conference abstract

#### 3.5.1 Panel A: the serosurveillance figure and RG/Bayesian adjustment
Why: they only alow on figure.
```{r}
## Load Yuan's RG data
rg_all <- read.csv("../1_data/private/df_time_rg_all.csv", row.names=1)
rg_ab  <- rg_all[rg_all$region=="Alberta" &
                 rg_all$source %in% c("APL", "CBS") &
                 !(rg_all$model=="RAW"),]
rg_ab$yearmon <- as.yearmon(rg_ab$yearmon)

## merge rg_ab and tb_alberta
tb_alberta<-merge(tb_alberta, select(rg_ab, 2:6),
      by.x = c("data", "samplemonth"), by.y = c("source", "yearmon"),
      all = T)   # need to ask Yuan about the updated CBS data

tb_ab_rg<-tb_alberta %>%
  select(-c("assay", "study")) %>%
  .[.$raw=="raw" & !is.na(.$raw),] %>%
  select(-c("raw"))

tb_ab_rg<-reshape2::melt(tb_ab_rg, 
                      id.vars = c("data", "samplemonth",
                                  "rg_lower", "rg_upper"),
                      variable.name = "rg",
                      value.name = "prop")
tb_ab_rg$rg_lower[tb_ab_rg$rg=="prop"]<-NA
tb_ab_rg$rg_upper[tb_ab_rg$rg=="prop"]<-NA
```

```{r}
# rebuild the figure
p03.5lab<-c("APL (raw)", "CBS (raw)", 
            "APL (RG-Bayesian)", "CBS (RG-Bayesian)")

p03.5<-tb_ab_rg %>%
  ggplot(aes(x=samplemonth, y=prop, colour=interaction(data,rg),
                                    linetype=interaction(data,rg))) +
  geom_point() +
  geom_line()  +
  geom_errorbar(aes(ymin=rg_lower, ymax=rg_upper), width=0.05)+
  theme_bw() + 
  xlab("Time of Sample Collection") + 
  ylab("Sero-prevalence") +
  scale_colour_manual("", values = c("red", "blue", "red", "blue"),
                      labels = p03.5lab) +
  scale_linetype_manual("", values=c(1,1,3,3), labels = p03.5lab) +
  theme(legend.position="top")+
  guides(color = guide_legend(nrow = 2))

p03.5
```

#### 3.5.2 Panel B: transformed assay result fig
```{r}
p03.3
```

#### 3.5.3 Panel C: rolling incidence fig
